% !TEX root = ../../math6370.tex
\newpage
\section{The Ring of Integers}
\subsection{Number Fields}

Algebraic Number Theory studies the arithmetic of number fields --- the ring of integers of number fields and their ideals, units, and factorizations. Number theoretic questions are then posed in terms of these algebraic objects and their properties. To begin, we recall the definition of the basic object of study in Algebraic Number Theory --- number fields. 


\begin{dfn}[Number Field]
A number field $K$ is a finite extension of $\Q$, i.e. a finite dimensional $\Q$-vector space. We can denote this pictorially as in the figure below. 
	\[
	\begin{tikzcd}
	K \arrow[dash]{d}{[K\,:\,\Q]=\,\dim_\Q K} \\
	\Q
	\end{tikzcd}
	\]
\end{dfn}


\begin{ex}[Number Fields] \label{ex:number_fields} \hfill
	\begin{enumerate}[(i)]
	\item The fields $\Q$, $\Q(\sqrt{2})$, $\Q(\sqrt[3]{2})$ are all number fields. These have dimension (as $\Q$-vector spaces) of 1, 2, and 3, respectively. 
	\item The cyclotomic fields: $\Q(\zeta_n)$, where $\zeta_n$ is a primitive $n$th root of unity.
	\item The field $\Q(\alpha)$, where $\alpha$ is a root of $x^3+2x^2+1$.
	\end{enumerate}
\end{ex}


Recall that a simple extension of a field $K$, say $L$, is an extension of the form $L= K(\alpha)$, where $\alpha$ is the root of some polynomial $p_\alpha(x) \in K[x]$. The examples of number fields in Example~\ref{ex:number_fields} are all simple extensions of $\Q$. These were not arbitrarily contrived `simple' examples of number fields. In fact, every number field arises in this way. Recall the following terminology: given a field extension $E/F$, an element $\alpha \in E$ is called a primitive element for $E/F$ if $E=F(\alpha)$. Then our preceding claim is that every number field possess a primitive element, which the following theorem justifies. 


\begin{thm}[Primitive Element Theorem] \label{thm:primelmthm}
If $K$ is a number field, then $K=\Q(\alpha)$, where $\alpha$ is the root of some polynomial $p_\alpha(x) \in \Q[x]$. 
\end{thm}

\pf See \cite[Ch. 14.4, Thm 25]{dummit}. Though there are proofs which avoid Galois Theory. For instance, the  proof of van der Waerden avoid Galois Theory and is `constructive.' This is essentially the proof given in \cite{murty}, but can also be found more thoroughly written out in \cite{brown}. \qed \pskip


Generally, if $E/F$ is a separable extension of finite degree, then $E=F(\alpha)$ for some $\alpha \in E$. Of course, the definition of a number field excludes infinite extensions of $\Q$ such as $\R$, $\C$, and $\Q(\{\sqrt[n]{2} \colon n \in \N\})$. Of course, there are even examples of infinite extensions of $\Q$ when the extension is simple, e.g. $K=\Q(e)$ or $K=\Q(\pi)$. But remember, the term `number field' will always refer to finite extensions of $\Q$. 


The traditional object of interest in number theory is $\Z \subseteq \Q$. What is `interesting' about $\Z$, among other things, is factorization. Of course, there is no factorization in $\Q$ because $\Q$ is a field. Of course, fields are `nicer' to work with. But constructing number fields, what should be the analog for a general number field $K$, i.e. what should the object $\O \subseteq K$ be so that factoring is `interesting'. What are the `integers' for a number field? The first goal for this course will be to define such an object and to study its properties. 


	\[
	\begin{tikzcd}
	? \arrow[dash]{r} & K \arrow[dash]{d}{[K \colon \Q]} \\
	\Z \arrow[dash]{r} & \Q
	\end{tikzcd}
	\]



% -------------------
% Ring of Integers
% -------------------
\subsection{The Ring of Integers}


Let $K$ be a number field. Choose $\alpha \in K$ and let $n=[K:\Q]$. Let $\phi: \Q[x] \to K$ be the homomorphism of $\Q$-algebras given by $x \mapsto \alpha$. This map is not injective (it is a map of an infinite dimensional vector space to a finite dimensional vector space). Therefore, the kernel is nonzero. The ring $\Q[x]$ is a PID so that the $\ker \phi$ is nonzero and principal. This induces an isomorphism $\widetilde{\phi}: \Q[x]/(p_\alpha(x)) \to K$, where $p_\alpha(x)$ is the minimal monic irreducible polynomial in $\Q[x]$ with $\alpha$ as a root (its so-called minimal polynomial). Denote by $\Q[\alpha]$ the image of $\widetilde{\phi}$ in $K$, i.e. the $\Q$-algebra generated by $\alpha$. Because $p_\alpha(x)$ is irreducible, $(p_\alpha(x))$ is maximal and hence $\Q[x]/(p_\alpha(x))$ is a field. But because $\Q[\alpha] \cong \Q[x]/(p_\alpha(x))$, we must have $\Q[\alpha] \cong \Q(\alpha)$ so that $\Q[\alpha]$ is a field. Note that since $\Q \subseteq K$, certainly $\Z \subsetneq K$. 


Now recall that $\alpha \in K/\Q$ is an algebraic number if $p_\alpha(x) \in \Q[x]$. [Such a $p_\alpha(x)$ exists because we cannot have $\alpha$ transcendental since we assume that $K$ is a number field.] So if $\alpha$ is an algebraic number, then the work in the preceding paragraph shows that $\Q[\alpha] \cong \Q(\alpha)$. Using this isomorphism, we shall often, by abuse of notation, treat $\Q[\alpha] = \Q(\alpha)$. We are interested in a very special type of algebraic number. 


\begin{dfn}[Algebraic Integer]
Let $K/\Q$ be a number field. We say $\alpha \in K$ is an algebraic integer if $p_\alpha(x)$ has coefficients in $\Z$, i.e. $p_\alpha(x) \in \Z[x]$. 
\end{dfn}


This definition allows us to define a candidate for a subring of $K$ where factorization will be `interesting.' 


\begin{dfn}[Ring of Integers]
Let $K/\Q$ be a number field. Define the ring of integers of  $K$, denoted $\O_K$, to be the set of algebraic integers of $K$. This is also sometimes denoted $\Z_K$. 
\end{dfn}


Despite being part of the definition, it is not immediately clear that $\O_K$ is even a ring. We shall prove this later. We will first look at some examples of rings of integers. The first example will show that this definition of `rings of integers' does reproduce the literal ring of integers $\Z$ in the case where $K= \Q$. [Note that because $p_r(x)= x - r \in \Z[x]$ for all $r \in \Z$, we always have $\Z \subseteq \O_K$.]


\begin{ex}\label{ex:zinq}
Take $K= \Q$. If $\alpha \in \Q$, then its minimal polynomial is $p_\alpha(x)= x - \alpha$. Now $p_\alpha(x) \in \Z[x]$ if and only if $\alpha \in \Z$. Therefore, $\O_\Q= \Z$. Then our definition of the ring of integers does indeed reproduce the motivating example of $\Z \subseteq \Q$. \xqed
\end{ex}


\begin{ex}
Take $K=\Q(\sqrt{-2})= \{ a + b \sqrt{-2} \colon a,b \in \Q \} \subseteq \C$ and $\alpha= a + b\sqrt{-2} \in K$. We have two cases: \pskip

\noindent $b=0$: If $b=0$, then $a \in \Q$. But then by the work in Example~\ref{ex:zinq}, $\alpha \in \O_K$ if and only if $\alpha \in \Z$. \pskip

\noindent $b\neq 0$: If $b \neq 0$,  then $p_\alpha(x)$ factors as 
	\[
	p_\alpha(x)=(x-(a+b\sqrt{-2}))(x-(a-b\sqrt{-2}))=x^2-2ax+(a^2+2b^2).
	\]
Therefore, $\alpha \in \O_K$ if and only if $2a \in \Z$ and $a^2+2b^2 \in \Z$. Let $N= a^2 + 2b^2$. Because $2a \in \Z$, we have $a= \frac{r}{2}$ for some $r \in \Z$. Then we have $N= a^2 + 2b^2= \frac{r^2}{4} + 2b^2$. But then clearing denominators shows $r^2 + 8b^2= 4N$, which implies that $r$ cannot be odd. Therefore, $r= 2k$ for some $k \in \Z$, which shows that $a= \frac{r}{2}= \frac{2k}{2}= k \in \Z$. 

But then $2b^2= N - a^2 \in \Z$ so that $2b^2= s$ for some $s \in \Z$. Using the fact that $b \in \Q$, write $b= \frac{u}{v}$ and $\gcd(u, v)= 1$. Then we have $2u^2= sv^2$. If $p$ is a prime dividing $v$, then $p^2 \mid (sv^2)= 2u^2$. Because $\gcd(u,v)=1$, we must have $p^2 \mid 2$, which is impossible because $2$ is not a square in $\Z$. Therefore, no prime divides $v$ so that $v= \pm 1$. But then $b^2= \frac{u^2}{v^2}= u^2$, forcing $b= \pm u \in \Z$. Therefore, we must have $a, b \in \Z$. \pskip
	
This shows that $\O_K= \Z[\sqrt{-2}]= \{ a + b \sqrt{-2} \colon a,b \in \Z \}$. Furthermore, one can show $\Z[\sqrt{-2}]$ is a UFD, i.e. factorial. \xqed
\end{ex}


There is one important motivating question we have not yet addressed: why do we care about `integers' in number fields rather than just considering the ordinary integers in $\Q \subseteq K$. The answer is simple: factorizations. In these larger rings, we might have `larger' and more `interesting' factorizations that we could use to solve problems that are otherwise more difficult than over $\Z$. 


\begin{ex}\label{ex:z(-sqrt(2))}
Find all solutions $x,y \in \Z$ for $y^2=x^3-2$. After a bit of experimentation, one can see that $(3,\pm 5)$ is a solution. However, are these the only solutions? Are there infinitely many integer solutions? We choose to work in $\Z[\sqrt{-2}]$ for the `extra factorizations' that it offers. Note that we shall use the fact that $\Z[\sqrt{-2}]$ is a UFD and $\Z[\sqrt{-2}]^\times=\{\pm 1\}$ without proof. [The latter will be easily proven using norms later, while the first fact follows from the fact that $\Z[\sqrt{-2}]$ is a Euclidean domain using the same canonical proof that $\Z[i]$ is Euclidean. All Euclidean domains are PIDs which are in turn UFDs.]

In $\Z[\sqrt{-2}]$, we have the factorization 
	\[
	x^3=y^2+2=(y+\sqrt{-2})(y-\sqrt{-2}). 
	\]
Write $x=u_1 \pi_1^{e_1}\cdots \pi_r^{e_r}$, where each $\pi_i$ is irreducible, $e_i \geq 1$, and $\pi_i \neq \pm \pi_j$ for $i \neq j$. Then we have
	\[
	u_1^3 \pi_1^{3e_1}\cdots \pi_r^{3e_r}=(y+\sqrt{-2})(y-\sqrt{-2}). 
	\]
We claim that $y+\sqrt{-2}$ and $y-\sqrt{-2}$ are relatively prime: choose an irreducible dividing both, say $\pi$. Then $\pi \mid \big(( y + \sqrt{-2}) - (y - \sqrt{-2}) \big)= 2\sqrt{-2}= (\sqrt{-2})^3$. By unique factorization in $\Z[\sqrt{-2}]$, up to a unit $u$, we must have $\pi= u\sqrt{-2}$. But the only units in $\Z[\sqrt{-2}]$ are $\{\pm 1\}$. Because $\pi^3= (\sqrt{-2})^3$, we must have $\pi= \sqrt{-2}$. Now as $\pi \mid (y + \sqrt{-2})$, we have $y + \sqrt{-2} = \pi (a + b \sqrt{-2})= \sqrt{-2} (a + b\sqrt{-2})= -2b + a\sqrt{-2}$ for some $a, b \in \Z$. But then as $\pi \mid (y + \sqrt{-2})$, we have $y + \sqrt{-2}= \sqrt{-2}(a + b\sqrt{-2})= -2b + a\sqrt{-2}$ for some $a, b \in \Z$. Therefore, $y= -2b$, which implies $x^3= y^2 + 2= 4b^2+2 \equiv 2 \mod 4$, a contradiction as no cube has residue 2 mod 4. This shows that $y + \sqrt{-2}$ and $y - \sqrt{-2}$ are relatively prime.

Then $\pi_i^{3e_i}$ divides $y+\sqrt{-2}$ or $y-\sqrt{-2}$ (but not both) for each $1 \leq i \leq r$. Then $y+\sqrt{-2}= u \prod_{i \in \I} \pi_i^{3e_i}$ for some $\I \subseteq \{ 1, \ldots, r \}$ and $u \in \Z[\sqrt{-2}]^\times= \{ \pm 1 \}$. This implies 
	\[
	y + \sqrt{-2}= u \prod_{i \in \I} \pi_i^{3e_i}= \prod_{i \in \I} \pi_i^{3e_i}= \left( \prod_{i \in \I} \pi_i^{e_i} \right)^3= (a + b\sqrt{-2} )^3
	\]
 for some $a,b \in \Z$. Expanding yields, $y+\sqrt{-2}=(a^3-6ab^2)+(3a^2b-2b^3)\sqrt{-2}$. This forces $y= a^3 - 6ab^2$ and $1= 3a^2b - 2b^3= b(3a^2 - 2b^2)$.

We now have a restriction on the possibilities for $y$! This restriction that would not have been as easily obtained working only in $\Z$. Now as $b (3a^2 - 2b^2)= 1$ with $a, b \in \Z$, it must be that $b \in \{\pm 1\}$. If $b= 1$, we have $3a^2 - 2= 1$, which gives $a= \pm 1$. Using $a= \pm 1$ and $b=1$, we have solutions $(3, \pm 5)$. If $b= -1$, then $3a^2(-1)-2(-1)^3=1$, which implies $3a^2=1$, a contradiction to the irrationality of $\sqrt{3}$. Therefore, the only integer solutions to $y^2= x^2 + 3$ are $(3, \pm 5)$. This also shows that this equation has only finitely many rational solutions.\footnote{On the other hand, there are infinitely many \emph{rational solutions} to this equation. In fact, this is an elliptic curve defined over $\Q$ with rank 1.} \xqed
\end{ex}


One must take care when using the factorization approach of Example~\ref{ex:z(-sqrt(2))}. Consider the following (non-)example. 


\begin{nex}\label{ex:factornonex}
Find all $(x,y) \in \Z^2$ satsifying $y^2=x^3-61$. Using the factorization in $\Z[\sqrt{-61}]$, we have $x^3=y^2+61=(y+\sqrt{-61})(y-\sqrt{-61})$. Using the method of Example~\ref{ex:z(-sqrt(2))}, one can show that there are no such solutions. But $(5, \pm 8)$ are both clearly solutions! \xqed
\end{nex}


The problem in Example~\ref{ex:factornonex} is that $\Z[\sqrt{-61}]$ is not a UFD: $5^3=8^2+61=(8+\sqrt{-61})(8-\sqrt{-61})$, where $5$, $(8+\sqrt{-61})$, and $(8-\sqrt{-61})$ are all irreducible in $\Z[\sqrt{-61}]$. Thus, Example~\ref{ex:z(-sqrt(2))} and Non-Example~\ref{ex:factornonex} show that not just factorization but \emph{unique} factorization is essential.  How does one cope with loss of unique factorization? Kummer (in approximately 1846) said there should be a further factorization into ``ideal numbers'' in order to recover unique factorization. This lead Dedekind to define ideals of a ring. He later gave the correct notion of factorization instead using (prime) ideals rather than factorization with elements. We shall see that while elements may/may not factor uniquely in $\O_K$, every ideal of $\O_K$ factors uniquely as a product of prime ideals. So we shall see that in Non-Example~\ref{ex:factornonex}, while elements $5$, $8 + \sqrt{-61}$, and $8 - \sqrt{-61}$ may not factor uniquely into irreducible, the ideals generated by these elements will factor uniquely into a product of prime ideals $\p, \q$:
	\[
	(5)= \p \q, \quad (8 + \sqrt{-61})= \p^3, \quad (8 - \sqrt{-61})= \q^3. 
	\]


To work towards this longterm goal, we want to study the properties of $\O_K$ further. As a short list of intermediate goals, we want to\dots
	\begin{itemize}
	\item Show $\O_K$ is a ring.
	\item Study the structure of abelian group $(\O_K, +)$ and show $\O_K \cong \Z^{[K:\Q]}$.
	\item Study the structure of units $\O_K^\times$.
	\item Show $\O_K$ has unique factorization into prime ideals.
	\item Measure failure of unique factorization, leading to the class group.
	\item Study the primes of $\O_K$ and how these primes `split'. 
	\end{itemize}


We shall begin by showing that $\O_K$ is a ring. The proof, which will be rather simple, will make use of some extra facts about algebraic integers.


\begin{prop}\label{prop:algint}
If $\alpha \in K$ is an algebraic integer, then the following are equivalent:
	\begin{enumerate}[(i)]
	\item $p_\alpha(x) \in \Z[x]$
	\item $f(\alpha)= 0$ for some monic polynomial $f(x) \in \Z[x]$
	\item $\Z[\alpha]$ is a finitely generated $\Z$-module
	\item there exists a nonzero finitely generated subgroup $M \subseteq K$ such that $\alpha M  \subseteq M$
	\end{enumerate}
\end{prop}

\pf We prove (i)$\iff$(ii) $\Rightarrow$ (iii) $\Rightarrow$ (iv) $\Rightarrow$ (i). 
\begin{enumerate}
\item[(i)$\to$(ii):] This is immediate.

\item[(ii)$\to$(iii):] If $f(x)=x^n + a_1x^{n-1} + \cdots + a_n$ has $\alpha$ as a root, then we can express the numbers $\{\alpha^i\}_{i \geq n}$ in terms of lower powers of $\alpha$. For example because $f(\alpha)= 0$, we have $\alpha^n= -(a_1\alpha^{n-1} + \cdots + a_n)$ and $\alpha^{n+1}= -\alpha (a_1\alpha^{n-1} + \cdots + a_n)= -a_1\alpha^n - \cdots - a_n\alpha= -a_1 (a_1\alpha^{n-1} + \cdots + a_n) - \cdots - a_n \alpha$. More concretely, there is a surjection $\Z[x] \to \Z[\alpha]$ given by $x \mapsto \alpha$. This gives a surjection $\Z[x]/(f) \to \Z[\alpha]$. But $\Z[x]/(f)$ is generated by $1,x,\ldots,x^{d-1}$, where $d=\deg(f)$. Therefore, $\Z[\alpha]$ is finitely generated. 

\item[(iii)$\to$(iv):] Take $M= \Z[\alpha]$. Then $\alpha M \subseteq M$. 

\item[(iv)$\to$(ii):] We have $M= \oplus_{i=1}^r \Z \beta_i$ for some collection $\{ \beta_i \}_{i=1}^r$. But then for each $i$, we can write $\alpha \beta_i= \sum_{j=1}^r c_{ij} \beta_j$ for some integer valued matrix $C= (c_{ij})$, i.e.
	\[
	C \begin{pmatrix} \beta_1 \\ \vdots \\ \beta_r \end{pmatrix}= \alpha \begin{pmatrix} \beta_1 \\ \vdots \\ \beta_r \end{pmatrix}.
	\]
But then $(C - \alpha I)(\beta_1,\ldots, \beta_r)^T= 0$. By the Cayley-Hamilton Theorem, $C$ is a root of its characteristic polynomial; that is, $C$ is a root of $\widetilde{f}(x)= \det(C - xI) \in \Z[x]$,
	\[
	\begin{vmatrix}
	c_{11} - x & c_{12} & \cdots & c_{1n} \\
	c_{21} & c_{22} - x & \cdots & c_{2n} \\
	\vdots & \vdots & \ddots & \vdots \\
	c_{n1} & c_{n2} & \cdots & c_{nn} - x 
	\end{vmatrix}= 0
	\]
 But then $\widetilde{f}(\alpha)= 0$. Defining $f(x)= (-1)^r \widetilde{f}(x)$, we have $f(x)$ is a monic polynomial with $f(\alpha)= 0$. 

\item[(ii)$\to$(i):] Suppose that $f(\alpha)=0$. Because $f(\alpha)= 0$, we know that $p_\alpha(x)$ divides $f(x)$. Suppose that $f(x)= p_\alpha(x)g(x)$ for some $g(x) \in \Q[x]$. It must be that both $p_\alpha(x)$ and $g(x)$ are monic because $f(x)= p_\alpha(x) g(x)$ is monic. It only remains to show that $p_\alpha(x) \in \Z[x]$. We claim that $p_\alpha(x),g(x) \in \Z[x]$. [This is really just Gau\ss' Lemma but we shall show it directly here.] If not, then fix a prime $p$ dividing the denominator of a coefficient of $p_\alpha(x)$ or $g(x)$. Take $e,f \geq 0$ such that $p^ep_\alpha(x)$ and $p^fg(x)$ have integer coefficients with no $p$'s in the denominator and $e,f$ are minimal (in the sense that choosing a smaller $e,f$ will not result in an integer valued polynomial). We have
	\[
	p^{e+f} f(x) = (p^e p^f) [p_\alpha(x) g(x) ]= p^e p_\alpha(x) \cdot p^f g(x).
	\]
Now take the equality $p^{e+f} f(x)= p^e p_\alpha(x) \cdot p^f g(x)$ modulo $p$. The right side must be nonzero for otherwise every coefficient is divisible by $p$ and hence have a $p$ in common. But then $e'= e-1, f'= f-1$ would have sufficed to clear denominators, contradicting the minimality of $e, f$. Therefore, the right side is nonzero modulo $p$. But then $p^{e+f} f(x) \not\equiv 0 \mod p$. Because $f(x)$ is monic, the leading coefficient of $p^{e+f} f(x)$ is $p^{e+f}$. But as $p^{e+f} f(x) \not\equiv 0 \mod p$, it must be that $e+f= 0$, i.e. $e= -f$. Because $e,f \geq 0$, we must have $e= f= 0$. Therefore, there is no prime dividing the denominator of any coefficient of $p_\alpha(x)$ or $g(x)$. Therefore, $p_\alpha(x),g(x) \in \Z[x]$, as desired. \qed 
\end{enumerate}


Note then we could have taken any of the (equivalent) properties of Proposition~\ref{prop:algint} as our definition of an algebraic integer. From now on, we shall take any of the equivalent conditions above as our definition of an algebraic integer, using whichever is most convenient in the moment. We are now in a position to easily prove that $\O_K$ is a subring of $K$. 


\begin{prop}\label{prop:subring}
$\O$ is a subring of $K$.
\end{prop}

\pf Observe that $0,1 \in \O_K$ so that $\O_K$ is nonempty. In particular, $\Z \subseteq \O_K$. Let $\alpha, \beta \in \O_K$. By Proposition~\ref{prop:algint}, the $\Z$-modules $\Z[\alpha]$ and $\Z[\beta]$ are finitely generated, say by $1, \alpha, \alpha^2, \ldots, \alpha^{d-1}$ and $1, \beta, \beta^2, \ldots, \beta^{e-1}$, respectively. Then $\Z[\alpha, \beta]$ is finitely generated by $\{\alpha^i \beta^j\}_{\substack{0 \leq i < d \\ 0 \leq j < e}}$. But $\alpha \pm \beta, \alpha\beta \in \Z[\alpha, \beta]$ so that $(\alpha \pm \beta) \Z[\alpha, \beta]$ and $(\alpha \beta) \Z[\alpha, \beta]$ are contained in $\Z[\alpha, \beta]$. By Proposition~\ref{prop:algint}, $\alpha \pm \beta, \alpha \beta$ are algebraic integers and hence $\alpha \pm \beta, \alpha \beta \in \O_K$. \qed \pskip


\noindent \textbf{Challenge Exercise:} Prove Proposition~\ref{prop:subring} using only (i) and (ii) of Proposition~\ref{prop:algint}. \pskip


The following proposition is also easily proven. 


\begin{prop} \label{prop:power_push} 
Let $\Q \subseteq K \subseteq L$ be number fields. Then
\begin{enumerate}[(i)]
\item $\O_K$ is an integral domain.
\item $\O_L \cap K= \O_K$. In particular, $\O_L \cap \Q= \Z$.
\item For any $\alpha \in K$, there is an integer $d \in \Z_{\geq 1}$ such that $d \alpha \in \O_K$.
\item $\Frac \O_K= K$.
\end{enumerate}
\end{prop}

\pf \hfill
\begin{enumerate}[(i)]
\item This is immediate because $\O_K \subseteq K$.

\item $\alpha \in \O_L \cap K$ if and only if $p_\alpha(x) \in \Z[x]$ and $\alpha \in K$ if and only if $\alpha \in \O_K$. In particular, $\O_L \cap \Q= \O_\Q= \Z$.

\item Take any $\alpha \in K$ and $f(x)= x^n + c_1 x^{n-1} + \cdots + c_n \in \Q[x]$ with $f(\alpha)= 0$. For $d \geq 1$, $d \alpha$ is a root of $x^n + d c_1 x^{n-1} + \cdots + d^n c_n$. Choosing $d$ sufficiently large, we can also assure that the coefficients are in $\Z$. 

\item Let $F$ denote the field of fractions of $\O_K$, i.e. $F= \Frac \O_K$. Because $F$ is the smallest field containing $\O_K$, we must have $F \subseteq K$. Now let $\alpha \in K$. Because $K$ is a number field, $\alpha$ is algebraic. But then by (iii), there exists $d \in \Z$ such that $d \alpha \in \O_K \subseteq F$. As $d \in \Z \subseteq F$, we must have $\alpha= d^{-1}(d\alpha) \in F$ so that $K \subseteq F$. Therefore, $K= \Frac \O_K$, as desired. \qed 
\end{enumerate}


Proposition~\ref{prop:power_push}~(iii) is particularly useful and is often not stated explicitly in the literature. In particular, this part shows that any element of a number field is `not far' from being an algebraic integer --- in addition to giving a short proof of Proposition~\ref{prop:power_push}~(iv).



% -------------------
% Trace and Norm
% -------------------
\subsection{Trace and Norm}

Suppose that $K/\Q$ has degree $n$. For $\alpha \in K$, define $\mu_\alpha: K \to K$ via $x \mapsto \alpha x$. This is a $\Q$-linear map so, fixing a basis, we can represent $\mu_\alpha$ by a $n \times n$ matrix. 

\begin{dfn}[Norm]
Define $\Nm{K/\Q}: K \to \Q$ via $\alpha \mapsto \det(\mu_\alpha)$. 
\end{dfn}

\begin{dfn}[Trace]
Define $\Tr{K/\Q}: K \to \Q$ via $\alpha \mapsto \trace(\mu_\alpha)$. 
\end{dfn}

\begin{rem}
The definitions of norm and trace do not depend on the fact that $K$ is a number field, only that the field extension is finite. So given any fields $K,F$ with $K/F$ finite, we may define $\Nm{K/F}$ and $\Tr{K/F}$. 
\end{rem}

\begin{rem}
While choosing different a different basis for $K/\Q$ will change the matrix $\mu_\alpha$, the trace and determinant of resulting matrix are invariant. This is because the new basis is conjugate to the old one and trace/determinant are invariant under conjugation.
\end{rem}

\begin{ex}
Let $K=\Q(\sqrt{d})$ with $d \neq 1$ a squarefree integer. This is a degree 2 extension and $K$ has basis $\{1,\sqrt{d}\}$. Take $\alpha= a+b\sqrt{d} \in K$, where $a,b \in \Q$. We have
	\[
	\begin{split}
	\alpha \cdot 1&= a+b\sqrt{d} \\
	\alpha \cdot \sqrt{d}&= bd+a\sqrt{d}
	\end{split}
	\]
Then relative to this basis, we have
	\[
	\mu_\alpha= \begin{pmatrix} a & bd \\ b & a \end{pmatrix}
	\]
We then have $\Nm{K/\Q}=a^2-db^2$ and $\Tr{K/\Q}=2a$. \xqed
\end{ex}

\begin{ex}
Let $K=\Q(\omega)$, where $\omega=\sqrt[3]{2}$. The minimal polynomial of $\omega$ over $\Q$ is $p_\omega(x)=x^3-2$. Therefore, $K$ is a degree 3 extension of $\Q$ with possible $\Q$-basis being $\{1,\omega,\omega^2\}$. Then for $\alpha \in K$, we may write $a+b\omega+c\omega^2$ for $a,b,c \in \Q$. With respect to the chosen basis, we have
	\[
	[\mu_\alpha]= \begin{pmatrix} a & 2c & 2b \\ b & a & 2c \\ c & b & a \end{pmatrix}
	\]
	\[
	\begin{split}
	\Tr{K/\Q}(\alpha)&=\trace[\mu_\alpha]=3a \\
	\Nm{K/\Q}(\alpha)&=\det[\mu_\alpha]=a^3+2b^3-4c^3-6abc
	\end{split}
	\]
Now when does $\Nm{K/\Q}(\alpha)=\pm1$? Certainly when $\alpha=\pm 1$, $\Nm{K/\Q}(\alpha)=\pm 1$. But this is also true for $\alpha=\ep$, where $\ep=1+\omega+\omega^2$. Since the norm is multiplicative, every power of $\ep$ has norm 1. It turns out, $\O_K=\Z[\omega]$ and $\O_K^\times=\{\pm \ep^n \colon n \in \Z\}$. \xqed
\end{ex}

From the properties of determinants and traces, it is routine to verify

\begin{prop} \hfill
	\begin{itemize}
	\item $\Nm{K/\Q}(\alpha\beta)=\Nm{K/\Q}(\alpha) \Nm{K/\Q}(\beta)$
	\item $\Nm{K/\Q}(c)=c^n$ for $c \in \Q$
	\end{itemize}
In particular, $\Nm{K/\Q}: K^\times \to \Q^\times$ is a homomorphism of groups.
	\begin{itemize}
	\item $\Tr{K/\Q}: K \to \Q$ is $\Q$-linear
	\end{itemize}
\end{prop}

\begin{ex}[Pell's Equation, $n=2$] \label{ex:pell2}
Using the norm, we shall show that $x^2-2y^2=1$ has infinitely many solutions. We know $\Z[\sqrt{2}] \subseteq K:=\Q(\sqrt{2})$. Observe $\Nm{K/\Q}=x^2-2y^2$. There is a correspondence
	\[
	\{(a,b) \in \Z^2 \colon a^2-2b^2=1\} \iff \{\alpha=a+b\sqrt{d} \in \Z[\sqrt{2}] \colon \Nm{K/\Q}(\alpha)=1\}
	\] 
Let $G=\{\alpha \in \Z[\sqrt{2}] \colon \Nm{K/\Q}(\alpha)=1\}$. We claim that $G$ is a subgroup of $\Z[\sqrt{2}]^\times$.
	\begin{itemize}
	\item Clearly, $1 \in G$.
	\item If $\alpha, \beta \in G$, then $\Nm{K/\Q}(\alpha\beta)=\Nm{K/\Q}(\alpha)\Nm{K/\Q}(\beta)=1\cdot 1=1$ so that $\alpha\beta \in G$.
	\item If $\alpha=a+b\sqrt{2} \in G$, then $\alpha^{-1}=a-b\sqrt{2} \in G$ as $\alpha\alpha^{-1}=a^2-2b^2=1$.
	\end{itemize}
Now take $\ep=3+2\sqrt{2} \in G$. Since $\ep>1$, the powers of $\ep$ are all distinct. But then there are infinitely many elements of norm 1; hence, there are infinitely many solutions to $x^2-2y^2=1$. In fact, $G=\pm \langle \ep \rangle=\{ \pm \ep^n \colon n \in \Z\}$. However, $\Z[\sqrt{2}]^\times= \pm \langle 1+\sqrt{2}\rangle$ and $G$ is a subgroup of $\Z[\sqrt{2}]^\times$ of index 2. Note that $(1+\sqrt{2})^2=3+2\sqrt{2}=\ep$. \xqed
\end{ex}

\begin{ex} \label{ex:pellgen}
In general, Pell's equation ($x^2-dy^2=1$ with $d>1$ squarefree) has infinitely many solutions. However, the first nontrivial solution (ones different from (1,0)), $\ep$, can be large. For example given $x^2-1141y^2=1$, the corresponding group is 
	\[
	G=\{ \alpha \in \Z[\sqrt{1141}] \colon \Nm{K/\Q}(\alpha)=1\}= \pm \langle \ep \rangle,
	\]
where $\ep=1036782394157223963237125215+30693385322765657197397208\sqrt{1141}$. \xqed
\end{ex}


These are special cases of a more general theorem, which will come later (see Section~\ref{sec:unit} and Section~\ref{sec:pell}): 




\begin{restatable*}[Dirichlet's Unit Theorem]{thm}{unit} \label{thm:unit}
Let $K$ be a number field of degree $n$ with $r$ real embeddings and $s$ conjugate pairs of complex embeddings. Then the abelian group $\O_K^\times$ is a finitely generated abelian group with rank $r+s-1$ and $\O_K^\times \cong \mu_K \times \Z^{r+s-1}$, where $\mu_K$ are the roots of unity in $\O_K$. 
\end{restatable*}

\begin{ex}
For a quadratic number field $K=\Q(\sqrt{d})$, with $d>0$ square free, $r=2$ and $s=0$ so that $\O_K^\times$ has rank $r+s-1=1$. \xqed
\end{ex}

\begin{ex}
The theorem also gives that when $K=\Q(\sqrt{d})$, it will not generally be the case that $\O_K= \Z[\sqrt{d}]$. For instance, if $\alpha=\dfrac{-1+\sqrt{-3}}{2}$. Observe that $\alpha^3=1$ so that $\alpha$ is a root of $x^3-1$; therefore, $\alpha \in \Q(\sqrt{-3})$ is an algebraic integer. However, $\alpha \notin \Z[\sqrt{-3}]$. \xqed
\end{ex}


\subsection{Complex Embeddings: Norm \& Trace Continued}


In trying to study the norm and trace, it is natural to study the characteristic polynomial of $[\mu_\alpha]$, i.e. $\det(xI-\mu_\alpha) \in \Q[x]$, as
	\[
	\det(xI-\mu_\alpha)=x^n - \Tr{K/\Q}(\alpha)x^{n-1} + \cdots + (-1)^n \Nm{K/\Q}(\alpha).
	\]

Now $\mu_\alpha$ acts on $K$ as a $\Q$-linear map. Extending scalars, $\mu_\alpha$ acts on $K \otimes_\Q \C$. Choose a primitive element $\beta$ for $K$. Then we have
	\[
	K=\Q(\beta) \cong \Q[x]/ (p_\beta(x)),
	\]
where $p_\beta(x)$ is the minimal polynomial of $\beta$. [The isomorphism in the left direction is $x \mapsto \beta$ or assuming equality, $\beta=[x]$.] We may then write
	\[
	K \otimes_\Q \C= \C[x]/(p_\beta(x)).
	\]
This ring is almost certainly not a field, since $\beta_beta(x)$ factors into linear terms over $\C$ by the Fundamental Theorem of Algebra. It should then be clear that this is a field if and only if $p_\beta$ has degree 1 if and only if $\beta \in \Q$ if and only if $K=\Q$. Now write $p_\beta(x)=\prod_{i=1}^n (x-\beta_i)$, where $\beta_i \in \C$ are the distinct roots of $p_\beta(x)$ in $\C$. By the Chinese Remainder Theorem,
	\[
	K \otimes_\Q \C= \C[x]/(p_\beta(x)) \cong \prod_{i=1}^n \C[x]/(x-\beta_i) \cong \prod_{i=1}^n \C =\C^n
	\]
This isomorphism is canonical up to choice of ordering of the roots and $\beta$. Using this isomorphism, $\mu_\alpha$ acts on $K \otimes_\Q \C \cong \C^n$ via multiplication by $\alpha \otimes 1$. Consider the composite
	\[
	\sigma_i: K \ma{} K \otimes_\Q \C \ma{\cong} \C^n \ma{\pi_i} \C
	\]
where the first map is $\alpha \mapsto \alpha \otimes 1$. Each map $\sigma_i: K \hra \C$ is a nonzero homomorphism of fields, hence injective. Since the isomorphism is determined by its action on $\beta$ (since the map will fix $\Q$), each embedding is characterized by the fact that $\sigma_i(\beta)=\beta_i$. 

\begin{dfn}[Complex Embeddings]
The field homomorphisms $\sigma_1,\ldots,\sigma_n: K \hra \C$ are the complex embeddings of $K$ into $\C$.
\end{dfn}

\begin{rem}
Note that if $K/\Q$ is a number field of degree $n$, the work above shows there are always exactly $n$ such embeddings. 
\end{rem}

Considering simple tensors, an explicit isomorphism $K \otimes_\Q \C \to \C^n$ is given by $\alpha \otimes z \mapsto (z\sigma_1(\alpha),\ldots,z\sigma_n(\alpha))$, i.e. `scaling'. The action on $K \otimes_\Q \C \cong \C^n$ can then be given by the diagonal matrix
	\[
	\begin{pmatrix}
	\sigma_1(\alpha) & & \\
	& \ddots & \\
	& & \sigma_n(\alpha)
	\end{pmatrix}
	\]
From this characterization, the following becomes clear:

\begin{prop}
For a number field $K/\Q$ of degree $n$,
	\[
	\begin{split}
	\Nm{K/\Q}(\alpha)&= \prod_{i=1}^n \sigma_i(\alpha) \\
	\Tr{K/\Q}(\alpha)&= \sum_{i=1}^n \sigma_i(\alpha)
	\end{split}
	\]
where $\sigma_1,\ldots,\sigma_n: K \hra \C$ are the complex embeddings of $K$.
\end{prop}

\begin{ex}
Let $K=\Q(\sqrt{d})$ be a quadratic extension, where $d \neq 1$ is a squarefree integer. We assume $K \subseteq \C$ already for convenience. There are two embeddings $\sigma_1,\sigma_2: K \hra \C$:
	\[
	\begin{split}
	\sigma_1(a+b\sqrt{d})&= a+b\sqrt{d} \\
	\sigma_2(a+b\sqrt{d})&= a-b\sqrt{d}
	\end{split}
	\]
Then the norm and trace are
	\[
	\begin{split}
	\Nm{K/\Q}(a+b\sqrt{d})&= (a+b\sqrt{d})(a-b\sqrt{d})= a^2-db^2 \\
	\Tr{K/\Q}(a+b\sqrt{d})&= (a+b\sqrt{d}) + (a-b\sqrt{d})= 2a
	\end{split}
	\] \xqed
\end{ex}


We had wanted to study more the relation between trace/norm and the characteristic polynomial of $[\mu_\alpha]$, i.e. $\det(xI-\mu_\alpha) \in \Q[x]$. The work above shows
	\[
	\det(xI - \mu_\alpha)= \prod_{i=1}^n (x - \sigma_i(\alpha))
	\]
Notice that $\sigma_i(\alpha)$ is a root of $p_\alpha(x)$ since $p_\alpha(x)$ has rational coefficients and each $\sigma_i$ is a morphism of $\Q$-algebras fixing these coefficients:
	\[
	0=\sigma_i(0)=\sigma_i(p_\alpha(\alpha))=p_\alpha(\sigma_i(\alpha))
	\]
Therefore,
	\[
	\det(xI - \mu_\alpha)= \prod_{i=1}^n (x - \sigma_i(\alpha))= p_\alpha^m
	\]
for some $m$. But we know $\deg p_\alpha=[\Q(\alpha):\Q]$ and $\deg \det(xI - \mu_\alpha)=n=[K:\Q]$. Therefore, we must have $m=[K:\Q(\alpha)]$. This proves the following:

\begin{prop}\label{prop:mult}
Let $K/\Q$ be a number field and $\alpha \in K$. Let $\mu_\alpha: K \to K$ denote multiplication by $\alpha$. Then
	\[
	\det(xI - \mu_\alpha)= \prod_{i=1}^n (x - \sigma_i(\alpha))= p_\alpha(x)^{[K:\Q(\alpha)]}
	\]
where $\sigma_1,\ldots,\sigma_n: K \hra \C$ are the complex embeddings of $K$ into $\C$.
\end{prop}

\begin{cor}\label{cor:mult}
For $\alpha \in \O_K$, $\Nm{K/\Q}(\alpha)$ and $\Tr{K/\Q}(\alpha)$ are integers.
\end{cor}

\pf From Proposition~\ref{prop:mult}, 
	\[
	\det(xI - \mu_\alpha)= p_\alpha(x)^{[K:\Q(\alpha)]}.
	\]
Since $p_\alpha(x) \in \Z[x]$, all the coefficients on the left side must also be integers. But we have
	\[
	\det(xI-\mu_\alpha)=x^n - \Tr{K/\Q}(\alpha)x^{n-1} + \cdots + (-1)^n \Nm{K/\Q}(\alpha).
	\] \qed \\


\begin{ex}
Let $K=\Q(\sqrt{d})$, where $d\neq 1$ is a squarefree integer. We want to find $\O_K$. If $\alpha=a+b\sqrt{d} \in \O_K$, where $a,b \in \Q$, we know from Corollary~\ref{cor:mult} that $\Nm{K/\Q}(\alpha),\Tr{K/Q}(\alpha) \in \Z$. Hence, $a^2-db^2,2a \in \Z$. Multiplying $a^2-db^2$ by 4, we obtain $(2a)^2-d(2b)^2,2a \in \Z$. 

Therefore, $2\O_K \subseteq \Z[\sqrt{d}]=\{a+b\sqrt{d} \colon a,b \in \Z\}$. Then we have an inclusion of abelian groups
	\[
	\Z[\sqrt{d}] \subseteq \O_K \subseteq \dfrac{1}{2}\, \Z[\sqrt{d}]
	\]
The quotient $\frac{1}{2}\Z[\sqrt{d}]/\Z[\sqrt{d}]$ is a group of order 4 with coset representatives: 0, $\frac{1}{2}$, $\frac{\sqrt{d}}{2}$, and $\frac{1+\sqrt{d}}{2}$. 

In order to determine $\O_K$, we need to determine which of these representatives are algebraic integers. Clearly, $0 \in \O_K$ and $\frac{1}{2} \notin \O_K$. The minimal polynomial of $\frac{\sqrt{d}}{2}$ is $x^2 - \frac{d}{4}$---which is not in $\Z[x]$ as $d$ is square free. Hence, $\frac{\sqrt{d}}{4} \notin \O_K$. Finally, the minimal polynomial of $\frac{1+\sqrt{d}}{2}$ is
	\[
	\left(x - \dfrac{1+\sqrt{d}}{2}\right) \left(x - \dfrac{1-\sqrt{d}}{2}\right)= x^2-x+ \dfrac{1-d}{4}.
	\]
Then $\frac{1+\sqrt{d}}{2}$ has minimal polynomial $p_\alpha(x) \in \Z[x]$. [That is, $\frac{1+\sqrt{d}}{2} \in \O_K$ if and only if $d \equiv 1 \mod 4$.] Therefore,
	\[
	\O+K=
	\begin{cases}
	\Z[\sqrt{d}], & d \not\equiv 1 \mod 4 \\
	\Z\left[\frac{1+\sqrt{d}}{2}\right], & d \equiv 1 \mod 4.
	\end{cases}
	\]
\xqed
\end{ex}


In the previous example where $K=\Q(\sqrt{d})$, where $d \neq 1$ is a squarefree integer, we found
	\[
	\O+K=
	\begin{cases}
	\Z[\sqrt{d}], & d \not\equiv 1 \mod 4 \\
	\Z\left[\frac{1+\sqrt{d}}{2}\right], & d \equiv 1 \mod 4.
	\end{cases}
	\]
But what is the point of this? Certainly, $\Z[\sqrt{d}] \subseteq \O_K$. Why not just use $\Z[\sqrt{d}]$? Why is $\O_K$ the `right' thing to study? As a vague answer, $\O_K$ is `nicer'. For example, take the case where $K=\Q(\sqrt{-3})$. We know $\O_K=\Z\left[\frac{1+\sqrt{-3}}{2}\right]$ not $\Z[\sqrt{-3}]$. Why use $\Z\left[\frac{1+\sqrt{-3}}{2}\right]$? Well, $\Z\left[\frac{1+\sqrt{-3}}{2}\right]$ is a UFD whereas $\Z[\sqrt{-3}] \subseteq \Z\left[\frac{1+\sqrt{-3}}{2}\right]$ is not. So the ring $\Z\left[\frac{1+\sqrt{-3}}{2}\right]$ is a `better' choice. There are under lying geometric considerations/connections here as well; the affine coordinate ring of $\Z\left[\frac{1+\sqrt{-3}}{2}\right]$ is smooth whereas the one for $\Z[\sqrt{-3}]$ is not. But for now, let us return to relations between $\O_K$ and $\Nm{K/\Q}$/$\Tr{K/\Q}$. 

\begin{prop}
$\alpha \in \O$ is a unit if and only if $\Nm{K/\Q}(\alpha)=\pm 1$.
\end{prop}

\pf ($\Longrightarrow$). If $\alpha$ is a unit, there is some $\beta \in K$ such that $\alpha \beta=1$. We have $\Nm{K/\Q}(\alpha)\Nm{K/\Q}(\beta)=\Nm{K/\Q}(\alpha\beta)=\Nm{K/\Q}(1)=1$. Since $\Nm{K/\Q}(\alpha)$ and $\Nm{K/\Q}(\beta)$ are integers, we must have $\Nm{K/\Q}(\alpha)=\pm 1$. 

($\Longleftarrow$). Identify $K \subseteq \C$ and let $\sigma_1,\ldots,\sigma_n: K \hra \C$ be the complex embeddings of $K$ into $\C$, relabeling if necessary, so that $\sigma_1=\text{id}$. Then
	\[
	\pm 1=\Nm{K/\Q}(\alpha)=\sigma_1(\alpha)\sigma_2(\alpha) \cdots \sigma_n(\alpha)= \alpha \cdot \sigma_2(\alpha) \cdots \sigma_n(\alpha).
	\]
Define $\beta:=\sigma_2(\alpha) \cdots \sigma_n(\alpha)$. Now $\beta \in K$ as $\alpha \in K$. We only need check that $\beta \in \O_K$. Now for all $i$, $\sigma_i(\alpha)$ is an algebraic integer since these are roots of the minimal polynomial $p_\alpha(x) \in \Z[x]$. But then $\beta:=\sigma_2(\alpha) \cdots \sigma_n(\alpha)$ is a product of algebraic integers so that $\beta \in \O_K$. Then $\alpha\beta=1$ with $\beta \in \O_K$. \qed 


\subsection{Integral Basis \& Orders}


We wanted to understand the additive abelian group structure of $\O_K$. Our next goal will be to prove that as an abelian group, $\O_K \cong \Z x_1 \oplus \cdots \oplus \Z x_n \cong \Z^n$. That is, $\{x_1,\ldots,x_n\}$ is an integral basis for $\O_K$. 

\begin{dfn}[Integral Basis]
A set of elements $\{x_1,\ldots,x_n\}$, where $x_i \in \O_K$ is called an integral basis for $\O_K$ if $\O_K \cong \Z x_1 \oplus \cdots \oplus \Z x_n \cong \Z^n$. We also say that this is an integral basis of $K$.
\end{dfn}

Any integral basis is also a basis of $K$ as a $\Q$-vector space. The image of these basis elements under the map
	\[
	\begin{tikzcd}
	K \arrow{r} & \displaystyle\prod_\sigma \, \C \\[-3ex]
	\alpha \arrow[maps to]{r} & (\sigma_(\alpha))_\sigma
	\end{tikzcd}
	\]
will lie in a smaller $\R$-vector space, which we shall define.

For $\delta: K \to \C$, denote by $\ov{\sigma}: K \hra \C$ the embedding obtained by composing $\sigma$ with complex conjugation. For $\alpha \in K$, we have $\ov{\sigma(\alpha)}= \ov{\sigma}(\alpha)$, essentially by definition. 

\begin{dfn}
$K_\R:= \{(\alpha_\sigma)_\sigma \in \prod_\sigma \C \colon \ov{\alpha}_\sigma = \alpha_{\ov{\sigma}}\}$.
\end{dfn}

There is a map
	\[
	\begin{tikzcd}
	K \arrow[hook]{r} & K_\R \\[-3ex]
	\alpha \arrow[maps to]{r} & (\sigma_(\alpha))_\sigma
	\end{tikzcd}
	\]

But what is the dimension, as an $\R$-vector space, of $K_\R$? We know $K \otimes_\Q \R \cong K_\R$ and $\dim_\R K_\R= \frac{1}{2} \dim_\R \left(\prod_\sigma \C\right)$, since if $\sigma=\ov{\sigma}$, then $\alpha_\sigma \in \R$, and if $\sigma \neq \ov{\sigma}$, then $\ov{\alpha}_\sigma= \alpha_{\ov{\sigma}}$, so knowing $\alpha_\sigma$ is enough to determine $\alpha_{\ov{\sigma}}$. [This is because $\ov{\alpha}_\sigma= \alpha_{\ov{\sigma}}$ if and only if $\real(\alpha_\sigma)=\real(\alpha_\sigma)$ and $\imag(\alpha_\sigma)= -\imag(\alpha_{\ov{\sigma}})$.] Then
	\[
	\dim_\R K_\R = \frac{1}{2} \, \dim_\R\left(\prod_\sigma \C \right) = \frac{1}{2} \cdot 2n=n
	\]

\begin{dfn}[Discrete]
An additive subgroup $H$ of $\C^n$ (or $\R^n$) is discrete (points are isolated) if $H \cap X$ is finite for any compact $X \subseteq \C^n$ (respectively, $X \subseteq \R^n$). 
\end{dfn}

\begin{ex} \hfill
	\begin{enumerate}[(i)]
	\item $\Z \subseteq \R \subseteq \C$ is discrete.
	\item $\Z\left[\frac{1+\sqrt{-3}}{2}\right] \subseteq \C$ is discrete.
	\item $\R \subseteq \C$ is not discrete. \xqed
	\end{enumerate} 
\end{ex}

\begin{ex}
Let $K=\Q(\sqrt{2}) \subseteq \R$. Then $\O_K=\Z[\sqrt{2}]$. The two embeddings $K \hra \C$ are
	\[
	\begin{split}
	\sigma_1(a+b\sqrt{2})&= a + b\sqrt{2} \\
	\sigma_2(a+b\sqrt{2})&= a - b\sqrt{2}
	\end{split}
	\]
These embeddings are self-conjugate so the map $K \hra K_\R=\R^2$ is given by $\alpha \mapsto (\sigma_1(\alpha),\sigma_2(\alpha))$. Note that $\Z[\sqrt{2}]$ is not discrete in $\R$ but considering $\Z[\sqrt{2}] \subseteq \R^2$ then $\O_K=\Z[\sqrt{2}]$ is discrete.
	\[
	\begin{tikzcd}
	\Z[\sqrt{2}] \arrow[hook]{r} & \R^2 \\[-3ex]
	a+b \sqrt{2} \arrow[mapsto]{r} & (a+b\sqrt{2},a-b\sqrt{2})
	\end{tikzcd}
	\]
We can picture this embedding as a lattice in $\R^2$ generated by vectors $\langle 1,-1\rangle$ and $\langle \sqrt{2}, -\sqrt{2}\rangle$. 

	\begin{figure}[h!]
	\centering
	\begin{tikzpicture}
	\clip (-3,-3) rectangle (3cm,3cm);
    		\coordinate (Origin)   at (0,0);
    		\coordinate (XAxisMin) at (-3,0);
    		\coordinate (XAxisMax) at (3,0);
    		\coordinate (YAxisMin) at (0,-3);
    		\coordinate (YAxisMax) at (0,3);
    		\draw [thin,-latex] (XAxisMin) -- (XAxisMax);
    		\draw [thin,-latex] (YAxisMin) -- (YAxisMax);
		
		\coordinate (VecOne) at (0.5,0.5);
		\coordinate (VecTwo) at (0.75,-0.75);
    		\draw [ultra thick,-latex,red] (Origin)
        			-- (VecOne) node [above left] {};
    		\draw [ultra thick,-latex,red] (Origin)
        			-- (VecTwo) node [below right] {};
			
    		\pgftransformcm{0.707107}{-0.707107}{0.707107}{0.707107}{\pgfpoint{0cm}{0cm}}	
		\foreach \x in {-5,...,5}	
    		\draw[style=help lines,dashed] ({1.0606*\x},{-3}) -- ({1.0606*\x},{3});
		\foreach \y in {-5,...,5}
    		\draw[style=help lines,dashed] ({-3},{0.7071*\y}) -- ({3},{0.7071*\y});
	\end{tikzpicture}
  	\caption{The lattice for $\Z[\sqrt{2}]$ generated by $\langle 1,1 \rangle$ and $\langle \sqrt{2}, -\sqrt{2}\rangle$. \label{fig:lattice}}
	\end{figure} \xqed
\end{ex}


This last example is not an aberration.

\begin{prop}
Under the inclusion $\iota: K \hra K_\R$, $\iota(\O_K)$ is a discrete subgroup of the $\R$-vector space $K_\R$; that is, all its points are isolated. 
\end{prop}

\pf Take $r>0$. It suffices to show there are only finitely many $\alpha \in \O_K$ with $|\sigma(\alpha)| \leq r$ for all $\sigma: K \hra \C$. Suppose $\alpha \in \O_K$ has this property. Then
	\[
	\prod_\sigma (x - \sigma(\alpha))= p_\alpha(x)^{[K:\Q(\alpha)]} \in \Z[x],
	\]
as $\alpha \in \O_K$. The coefficients of the polynomial on the left are bounded, by the assumption that $|\sigma(\alpha)| \leq r$, in terms of $r$ and the degree $n$. Hence, there are only finitely many possibilities for $\prod_\sigma (x - \sigma(\alpha))$ since $\Z \subseteq \R$ is discrete. Hence, there are only finitely many such $\alpha$. \qed \\


While the next proposition is not directly related to the topics at hand, it will be useful later.

\begin{prop}\label{prop:zbasis}
Let $H$ be a discrete subgroup of $\R^n$. Then $H$ is a free $\Z$-module of rank at most $n$. Then any $\Z$-basis of $H$ is linearly independent over $\R$. 
\end{prop}

\pf Let $V:= \Span_\R H$. Choose a basis for $V$ in $H$. We may assume $\Z^r \subseteq H \subseteq \R^r$, where $r=\dim_\R V \leq n$. Replacing $\R^n$ by $V$ if necessary, we may assume $\Z^n \subseteq H \subseteq \R^n$. 

Now every coset of $H/\Z^n$ has a representative with $(a_1,\ldots,a_n) \in \R^n$ with $0 \leq a_i<1$ (simply by rounding down each component and subtracting off a vector consisting of these numbers). Since $H$ is discrete, there are only finitely many such representatives. Hence, the quotient is finite.

Set $m:= \#(H/\Z^n)$. Multiplying any element of $H/\Z^n$ by $m$ must give the identity coset, so multiplying any element of $H$ by $m$ lands in $\Z^n$. So $\Z^n \subseteq H \subseteq \frac{1}{m} \Z^n$. Therefore, $H$ must be free of rank $n$ as a $\Z$-module, i.e $H \cong \Z^n$. \qed \\

We can then obtain one of our long term goals, describing $(\O_K,+)$.

\begin{prop}\label{prop:ringintabelian}
Let $K/\Q$ be a number field. Then $\O_K \cong \Z^n$ as an additive abelian group.
\end{prop}

\pf By Proposition~\ref{prop:zbasis}, we know that $\O_K \cong \iota(\O_K) \subseteq K_\R \cong \R^n$. Therefore, $\O_K \cong \Z^r$, where $r \leq n$. Now take a basis $\{x_1,\ldots,x_n\}$ of $K$ over $\Q$. After multiplication by some integer $m \geq 1$, i.e. clearing denominators (this does not affect linear independence), we may assume that $x_i \in \O_K$. Then $x_1,\ldots,x_n$ are linearly independent elements in $(\O_K,+)$. Therefore, $r \geq n$. But then we must have $r=n$. Hence, $\O_K \cong \Z^n$ as an additive abelian group. \qed \\




In general, how does one compute $\O_K$? We know how to compute $\O_K$ for quadratic extensions but very little else. We will introduce a `good guess' for the ring of integers (this will be an order). For instance, if $K=\Q(\sqrt[3]{2})$, one would `guess' $\O_K \stackrel{?}{=} \Z[\sqrt[3]{2}]$. The theory of discriminants will then allow us to check and adjust our guess, allowing us to work our way to the true ring of integers. 

\begin{dfn}[Order]
Let $K$ be a number field. An order of $K$ is a subring $R \subseteq K$ that is isomorphic as an additive group to $\Z^n$, where $n=[K:\Q]$. 
\end{dfn}

\begin{ex}
If $K=\Q(\sqrt[3]{2})$, then $\Z[\sqrt[3]{2}]$ is an order of $K$. \xqed
\end{ex}

\begin{ex}
By Proposition~\ref{prop:ringintabelian}, $\O_K$ is always an order of $K$. \xqed
\end{ex}

Indeed, $\O_K$ is the maximal order of $K$. One can define $\O_K$ to be the maximal order of $K$. However, it does take a bit of work to then show it is equivalent to our definition. 

\begin{lem}
For any order $R \subseteq K$, $R \subseteq \O_K$, i.e. $\O_K$ is the maximal order of $K$.
\end{lem}

\pf Take $\alpha \in R$. Then $\alpha R \subseteq R$. Note that $R$ is an order, i.e. a finitely generated $\Z$-submodule of $K$. But then by Propostion~\ref{prop:algint}, we know that $\alpha \in \O_K$. \qed \\

Now we have a symmetric (non-degenerate) $\Q$-bilinear pairing on $K$:
	\[
	\begin{tikzcd}
	\langle \;,\; \rangle \arrow{r} & \Q \\[-3ex]
	(\alpha,\beta) \arrow[mapsto]{r} & \Tr{K/\Q}(\alpha\beta)
	\end{tikzcd}
	\]
For any order $R$, $\langle \;,\;\rangle: R \times R \to \Z$ since $(\alpha,\beta) \mapsto \Tr{K/\Q}(\alpha\beta) \in \Z$ by Corollary~\ref{cor:mult}. 

\begin{rem}
This induces a pairing on $K \otimes_\Q \R \cong K_\R$ and gives it a Euclidean structure, i.e. a standard inner product on $\R^n$.
\end{rem}

Fix a basis $\{x_1,\ldots,x_r\}$ of $R$ as a $\Z$-module, and take $\alpha,\beta \in R$, writing them as
	\[
	\begin{split}
	\alpha&= \sum_{i=1}^r a_i x_i \\
	\beta&= \sum_{i=1}^r b_i x_i,
	\end{split}
	\]
where $a_i,b_i \in \Z$. Expanding using linearity, 
	\[
	\Tr{K/\Q}(\alpha\beta)= \sum_{i,j} a_i \Tr{K/\Q}(x_ix_j)b_j= (a_1,\ldots,a_n) 
	\begin{pmatrix}
	\Tr{K/\Q}(x_1x_1) & \cdots & \Tr{K/\Q}(x_1x_r) \\
	\vdots & \ddots & \vdots \\
	\Tr{K/\Q}(x_rx_1) & \cdots & \Tr{K/\Q}(x_rx_r)
	\end{pmatrix}
	\begin{pmatrix}
	b_1 \\ \vdots \\ b_r
	\end{pmatrix}
	\]

\begin{dfn}[Discriminant]
The discriminant of an $n$-tuple $(x_1,\ldots,x_n) \in \O_K$ is
	\[
	\disc(x_1,\ldots,x_n):= \det\left(\Tr{K/\Q}(x_ix_j)\right)_{i,j}
	\]
\end{dfn}

Note that the discriminant takes values in $\Z$. But how does this depend on the choice of basis? If one chose another basis of $R$, say $\{y_1,\ldots,y_n\}$, write
	\[
	y_i = \sum_{j=1}^r B_{ij} x_j
	\]
for some $B_{ij} \in \Z$. Define the matrix $B=(B_{ij})_{ij}$. The matrix $B$ is invertible, i.e. $B \in \GL_n(\Z)$ and clearly symmetric. It is routine to verify that
	\[
	\left( \Tr{K/\Q}(y_iy_j) \right)_{ij}= B \left(\Tr{K/\Q}(x_ix_j) \right)_{ij} B^T
	\]
Then taking determinants, we have
	\[
	\disc(y_1,\ldots,y_n)= \det(B)^2 \disc(x_1,\ldots,x_n),
	\]
but since $B \in \GL(\Z)$, $\det(B)=\pm 1$. Therefore, the discriminant is independent of the choice of basis. 

\begin{dfn}[Discriminant (Order)]
The discriminant of an order $R$ is $\disc(R):=\disc(x_1,\ldots,x_n)$, where $\{x_1,\ldots,x_n\}$ is a basis of $R$.
\end{dfn}

We can now make our previous `guess-and-check' notion for calculating $\O_K$ a bit more precise. For an order $R \subseteq \O_K$, $\disc(R)=\disc(\O_K) [\O_K \colon R]^2$ (as we shall show). Then to find $\O_K$:
	\begin{enumerate}[1.]
	\item Guess an order $R$.
	\item Compute $\disc(R)$, which gives a finite number of possibilities for $[\O_K \colon R]$.
	\item Observe $R \subseteq \O_K \subseteq \frac{1}{m} R$. Find representatives for the cosets and determine which are algebraic integers. 
	\end{enumerate}

\begin{prop}
For an order $R \subseteq K$, $\disc(R) \neq 0$.
\end{prop}

\pf Suppose $0 \neq \alpha \in R$. Then $\Tr{K/\Q}(\alpha\alpha^{-1})=[K \colon \Q] \neq 0$. There is some $m \geq 1$ such that $m \alpha^{-1} \in R$. Then $\langle \alpha,\beta \rangle= m[K \colon \Q] \neq 0$ since $\langle \;,\; \rangle$ is non-degenerate. \qed 


\subsection{Discriminants}


\begin{dfn}[Discriminant (Number Field)]
If $K$ is a number field, then we define $\disc(K):= \disc(\O_K)$. 
\end{dfn}

\begin{rem}
There are many notations for what we have denoted as $\disc(K)$: $d_K,D_K,\Delta_K,\ldots$
\end{rem}


Before we prove a useful lemma relating $\disc(R)$ and $\disc(\O_K)$, we shall remind the reader of Smith Normal Form.

\begin{thm}[Smith Normal Form]
Let $B \in M_n(\Z)$ be of rank $r$. There exists matrices $P,Q \in \GL(\Z)$ such that
	\[
	PBQ=
	\begin{pmatrix}
	 d_1 &  &  &  &  & \\ 
	 &  \ddots &  &  &  & \\ 
	 &  &  d_r &  &  & \\ 
	 &  &  & 0 &  & \\ 
	 &  &  &  & \ddots & \\ 
	 &  &  &  &  & 0
	\end{pmatrix},
	\]
where $d_i \geq 1$ and $d_i \mid d_{i+1}$. 
\end{thm}

\begin{rem}
One only need the entries of the matrix be in a PID and the matrix need not be square. One defines $d_i=\div_i(A)/\div_{i-1}(A)$, where $\div_i$ is the $i$th determinant divisor, i.e. the gcd of all the $i \times i$ minors of $A$ and $d_0(A):=1$. 
\end{rem}


\noindent \textbf{Exercise: }Prove the classification of finitely generated abelian groups using the Smith Normal Form. \\


\begin{lem}
Let $R \subseteq K$ be an order. Then $\disc(R)= \disc(\O_K) [\O_K \colon R]^2$. 
\end{lem}

\pf Write $\O_K= \Z x_1 \oplus \cdots \oplus \Z x_n$ and let $R \subseteq \O_K$ be an order. Write $R= \Z y_1 \oplus \cdots \oplus \Z y_n$. For each of the $y_i$'s, write
	\[
	y_i \sum_{j=1}^n B_{ij} x_i
	\]
Define $B:=(B_{ij})_{ij}$. We know that $B \in \GL_n(\Z)$. Observe that $[\O_K \colon R] = [\Z^n \colon B(\Z^n)]$. By some previous remarks,
	\[
	\disc(y_1,\ldots,y_n)= \det(B)^2 \disc(x_1,\ldots,x_n)
	\]
Therefore, $\disc(R)= \det(B)^2 \disc(\O_K)$. It remains to show that $\det(B)= \pm [\O_K \colon R]$. Without loss of generality, assume that $B$ is of the form
	\[
	B=\begin{pmatrix} d_1 & & \\ & \ddots & \\ & & d_n \end{pmatrix}
	\]
with $d_i \geq 1$ and $\det B \neq 0$. [Using Smith Normal Form, we replace $B$ by $PBQ$. This does not change $[\Z^n \colon B(\Z^n)]$ or $\pm\det(B)$ since $P,Q$ are invertible integer matrices. Then $\det(P),\det(Q) \in \{\pm 1\}$.] Now $\det(B)=d_1d_2\cdots d_n$. Therefore,
	\[
	B(\Z^n)= d_1 \Z \times \cdots \times d_n \Z
	\]
and
	\[
	\Z^n/B(\Z^n) \cong \Z/d_1\Z \times \cdots \times \Z/d_n\Z,
	\]
which has cardinality $d_1\cdots d_n$, as desired. \qed \\

\begin{ex}
Let $K=\Q(\sqrt{d})$, where $d \neq 1$ is a squarefree integer. Now $R=\Z[\sqrt{d}]=\Z + \Z \sqrt{d}$ is an order of $K$. Furthermore,
	\[
	\begin{split}
	\disc(R)&= \disc(1,\sqrt{d}) \\
	&= \det \begin{pmatrix}
	\Tr{K/\Q}(1 \cdot 1) & \Tr{K/\Q}(1 \cdot \sqrt{d}) \\
	\Tr{K/\Q}(\sqrt{d} \cdot 1) & \Tr{K/\Q}(\sqrt{d} \cdot \sqrt{d})
	\end{pmatrix} \\
	&= \det \begin{pmatrix} 2 & 0 \\ 0 & 2d \end{pmatrix} \\
	&=4d.
	\end{split}
	\]
So $4d=\disc(R)=\disc(\O_K) [\O_K \colon R]^2$. Since $d$ is squarefree, this implies that $[\O_K \colon R] \in \{1,2\}$; that is, $\O_K=R$ or is index 2 larger, i.e.	
	\[
	R \subseteq \O_K \subseteq \dfrac{1}{2}\, R.
	\]
Now $R/\frac{1}{2}R$ has coset representatives $1, \frac{1}{2}, \frac{\sqrt{d}}{2}$, and $\frac{1+\sqrt{d}}{2}$. One need only check which of these are algebraic integers. It is routine to check that 1 is an algebraic integer, $\frac{1}{2}, \frac{\sqrt{d}}{2}$ are not algebraic integers, and $\frac{1+\sqrt{d}}{2}$ is sometimes an algebraic integer (if and only if $d \equiv 1 \mod 4$). Therefore,
	\[
	\begin{split}
	\O_K&=
	\begin{cases}
	\Z[\sqrt{d}], & \text{if } d \not\equiv 1 \mod 4 \\
	\Z\left[\frac{1+\sqrt{d}}{2}\right], & \text{if } d \equiv 1 \mod 4
	\end{cases} \\
	\disc(\O_K)&=
	\begin{cases}
	4d, & \text{if } d \not\equiv 1 \mod 4 \\
	d, & \text{if } d \equiv 1 \mod 4.
	\end{cases}
	\end{split}
	\] \xqed
\end{ex}

\begin{rem}
The discriminant of $\O_K$ determines a degree 2 extension $K/\Q$ up to isomorphism. Later, we shall see that there are only finitely many number fields of any degree with a given discriminant. 
\end{rem}

\begin{lem}\label{lem:embeddisc}
Let $\sigma_1,\ldots,\sigma_n: K \hra \C$ be the complex embeddings of $K$ into $\C$. Then for $x_1,\ldots,x_n \in \O_K$,
	\[
	\disc(x_1,\ldots,x_n)= \bigg(\det \big(\sigma_i(x_j)\big)_{i,j} \bigg)^2= \left[\det 
	\begin{pmatrix}
	\sigma_1(x_1) & \cdots & \sigma_1(x_n) \\
	\vdots & \ddots & \vdots \\
	\sigma_n(x_1) & \cdots & \sigma_n(x_n)
	\end{pmatrix}\right]^2
	\]
\end{lem}

\pf Recall $\Tr{K/\Q}(x)= \sum_{i=1}^n \sigma_i(x)$. Let $A:=(\sigma_i(x_j))_{ij}$. Then we have
	\[
	\begin{split}
	\disc(x_1,\ldots,x_n)&= \det\bigg(\Tr{K/\Q}(x_ix_j)\bigg)_{i,j} \\
	&=\det\left(\sum_{k=1}^n \sigma_k(x_i) \sigma_k(x_j) \right) \\
	&=\det\left(\sum_{k=1}^n \sigma_k(x_i x_j) \right) \\
	&= \det( A^TA) \\
	&= \left[ \det(A) \right]^2.
	\end{split}
	\]
\qed \\


\begin{ex}
Let $K=\Q(\sqrt{d})$, where $d \neq 1$ is a squarefree integer. Consider $R=\Z[\sqrt{d}] \subseteq \Q(\sqrt{d})=K$. The two complex embeddings of $K$ into $\C$ are 
	\[
	\begin{split}
	\sigma_1(a+b\sqrt{d})&= a + b\sqrt{d} \\
	\sigma_2(a+b\sqrt{d})&= a - b\sqrt{d}
	\end{split}
	\]
Then by Lemma~\ref{lem:embeddisc}
	\[
	\begin{split}
	\disc R&= \disc(1,\sqrt{d}) \\
	&=\left[\det \begin{pmatrix}
	\sigma_1(1) & \sigma_1(\sqrt{d}) \\
	\sigma_2(1) & \sigma_2(\sqrt{d})
	\end{pmatrix} \right]^2 \\
	&= \left[\det \begin{pmatrix}
	1 & \sqrt{d} \\
	1 & -\sqrt{d}
	\end{pmatrix} \right]^2 \\
	&= 4d
	\end{split}
	\] \xqed
\end{ex}


The term `discriminant' will already be familiar to the reader from the discriminant of a polynomial. 

\begin{dfn}[Discriminant (Polynomial)]
If $f(x) \in \Q[x]$ is a monic polynomial of degree $n \geq 1$, the discriminant of $f$ is
	\[
	\disc(f)= \prod_{1 \leq i<j \leq n} (\alpha_i - \alpha_j)^2
	\]
where $\alpha_1,\ldots,\alpha_n$ are the roots of $f$ in $\C$. 
\end{dfn}

It is worth noting that $\disc(f) \in \Q$ and if $f \in \Z[x]$ then $\disc(f) \in \Z$. Note also that if the polynomial is not monic, one can always divide by the leading coefficient and use the fact that if $\alpha \in \Q$, then $\disc(\alpha x)= \alpha^n \disc(x)$, where $n=[K:\Q]$. 

\begin{ex}
Here are some discriminants of several specific polynomials:
	\[
	\begin{split}
	\disc(x^2+bc+x)&= b^2 - 4c \\
	\disc(x^3+bx^2+cx+d)&= b^2c^2 - 4c^3 - 4b^3d - 27d^2 + 18bcd \\
	\disc(x^3+cx+d)&= -4c^3 - 27 d^2
	\end{split}
	\] \xqed
\end{ex}

We shall connect the discriminant of a number field with that of the discriminant of a polynomial. 

\begin{lem}\label{lem:discmin}
If $\alpha \in \O_K$ is such that $K=\Q(\alpha)$ (so that $\Z[\alpha]$ is an order of $K$), then $\disc(\Z[\alpha])= \disc(p_\alpha(x))$, where $p_\alpha(x)$ is the minimal polynomial of $\alpha$ over $\Q$. 
\end{lem}

\pf Let $\sigma_1,\ldots,\sigma_n: K \hra \C$ be the complex embeddings of $K$ into $\C$. The roots of $p_\alpha(x)$ are $\sigma_1(\alpha),\ldots,\sigma_n(\alpha)$. Write $\alpha_i= \sigma_i(\alpha)$. The order $\Z[\alpha]$ has a  integral $\Z$-basis $\{1,\alpha,\ldots,\alpha^{n-1}\}$. Therefore,
	\[
	\begin{split}
	\disc(\Z[\alpha])&= \disc(1,\alpha,\ldots,\alpha^{n-1}) \\
	&= \left( \det \begin{pmatrix}
	1 & \alpha_1 & \alpha_1^2 & \cdots & \alpha_1^{n-1} \\
	1 & \alpha_2 & \alpha_2^2 & \cdots & \alpha_2^{n-1} \\
	\vdots & \vdots & \vdots^2 & \ddots & \vdots \\
	1 & \alpha_n & \alpha_n^2 & \cdots & \alpha_n^{n-1} 
	\end{pmatrix} \right)^2 \\
	&= \left( \prod_{1 \leq i < j \leq n} (\alpha_j - \alpha_i) \right)^2 \\
	&= \disc(p_\alpha(x))
	\end{split}
	\] \qed \\

\begin{ex}
Let $K=\Q(\alpha)$, where $\alpha$ is a root of $f(x)=x^3+x+1$. Observe that $f(x)$ is irreducible and then $[K:\Q]=3$. Now $\Z[\alpha]$ is an order of $K$ with 
	\[
	\disc(\Z[\alpha])= \disc(f)= -4(1)^3 - 27(1)^2= -31.
	\]
But $\disc(\Z[\alpha])= \disc(\O_K) [\O_K \colon \Z[\alpha]]^2$. Since 31 is prime, we must have $[\O_K \colon \Z[\alpha]=1$ which implies that $\O_K=\Z[\alpha]$ and $\disc(K)= -31$. \xqed
\end{ex}

\begin{ex}
Let $K=\Q(\alpha)$, where $\alpha$ is a root of $f(x)= x^3-x^2-2x-8$. Consider the order $\Z[\alpha]$ of $K$. 
	\[
	\disc(\Z[\alpha])= \disc(f)= -2012= -2^2 \cdot 503.
	\]
Then $[\O_K \colon \Z[\alpha]] \in \{1,2\}$. Therefore,
	\[
	\Z[\alpha] \subseteq \O_K \subseteq \dfrac{1}{2} \, \Z[\alpha]
	\]
The group $\frac{1}{2}\Z[\alpha]/\Z[\alpha$ has coset representatives $\frac{a}{2} + \frac{b}{2} \alpha + \frac{c}{2} \alpha^2$, where $a,b,c \in \{0,1\}$. We need only check which of these eight elements are algebraic integers. Let $\theta:= \frac{\alpha + \alpha^2}{2}$. Now $K$ has $\Q$-basis $\{1,\alpha,\alpha^2\}$ and $\theta$ acts on this by
	\[
	\begin{split}
	\theta \cdot 1&= \frac{1}{2} \alpha + \frac{1}{2} \alpha^2 \\
	\theta \cdot \alpha&= 4+\alpha + 2\alpha^2 \\
	\theta \cdot \alpha^2&= 8 + 6\alpha +2\alpha^2
	\end{split}
	\]
Therefore, $\theta$ is a root of 
	\[
	\det\left(xI - 
	\begin{pmatrix}
	0 & 4 & 8 \\
	^1/_2 & 1 & 6 \\
	^1/_2 & 1 & 2 
	\end{pmatrix}
	\right)= x^3 - 3x^2 -10x - 8.
	\]
But then $\theta \in \O_K$. Therefore, $\O_K= \Z 1 \oplus \Z \alpha \oplus \Z\frac{\alpha+\alpha^2}{2}=\Z \oplus \Z \alpha \oplus \Z \theta$. Note that $\O_K \neq \Z[\beta]$ for any $\beta \in \O_K$ --- a result due to Dedekind. \xqed
\end{ex}


\begin{ex}
In the previous example, we saw that for $K=\Q(\alpha)$, where $\alpha$ is a root of $f(x)=x^3-x^2-2x-8$, $\O_K \neq \Z[\beta]$ for any $\beta \in \O_K$. This is not the only example. For example, if $K=\Q(\sqrt[3]{19})$, then $\O_K \neq \Z[\sqrt[3]{19}]$ but rather $\O_K=\Z+\Z\omega+ \Z \frac{1+\omega+\omega^2}{3}$, where $\omega=\sqrt[3]{19}$. \xqed
\end{ex}

We have seen how the discriminant relates to $\O_K$. We shall now see how the discriminant relates to the norm.

\begin{lem}\label{lem:discnorm}
Let $K=\Q(\alpha)$ be a number field, where $\alpha \in \O_K$. If $f \in \Z[x]$ is the minimal polynomial of $\alpha$ over $\Q$, then	
	\[
	\disc(\Z[\alpha])= \disc(f)= (-1)^{\frac{n(n-1)}{2}} \Nm{K/\Q}(f'(\alpha))
	\]
In particular, $\disc(\Z[\alpha]) \in \Q$. 
\end{lem}

\pf The first equality was Lemma~\ref{lem:discmin}. For the second equality, let $\sigma_1,\ldots,\sigma_n: K \hra \C$ be the complex embeddings of $K$ into $\C$. By rearranging terms,
	\[
	\begin{split}
	\disc f&= \prod_{1 \leq i < j \leq n} (\sigma_i(\alpha) - \sigma_j(\alpha))^2 \\
	&= (-1)^{\binom{n}{2}} \prod_j \prod_{i \neq j} (\sigma_j(\alpha) - \sigma_i(\alpha))
	\end{split}
	\]
However, we know $f=\prod_i (x - \sigma_i(\alpha))$. Using the product rule, this gives 
	\[
	f'= \sum_j \prod_{i \neq j} (x - \sigma_i(\alpha)).
	\]
But then $f'(\sigma_j(\alpha))= \prod_{i \neq j} (\sigma_j(\alpha) - \sigma_i(\alpha))$ since all but one term vanishes. Using this in the above along with the fact that the $\sigma_i$ fix $\Q$ (hence fix the coefficients of $f'$), we have
	\[
	\begin{split}
	(-1)^{\binom{n}{2}} \prod_j \prod_{i \neq j} (\sigma_j(\alpha) - \sigma_i(\alpha))&= (-1)^{\frac{n(n-1)}{2}} \prod_j f'(\sigma_j(\alpha)) \\
	&= (-1)^{\frac{n(n-1)}{2}} \prod_j \sigma_j(f'(\alpha)) \\
	&=  (-1)^{\frac{n(n-1)}{2}} \Nm{K/\Q}(f'(\alpha))
	\end{split}
	\] \qed \\

\begin{ex}[$p$th Cyclotomic Field]
Fix an odd prime $p$. Let $\zeta \neq 1$ be a $p$th root of unity, i.e. $\zeta^p=1$. Let $K=\Q(\zeta)$ be the $p$th cyclotomic field. Define
	\[
	f(x):= \dfrac{x^p-1}{x-1}= x^{p-1} + x^{p-2} + \cdots + x + 1 \in \Z[x]
	\]
Observe that $f(\zeta)=0$ and that $f$ is irreducible (examine $f(x+1)$, use Eisenstein's criterion, then invoke Gau\ss' lemma). In particular, $[K\colon \Q]=p-1$. Now what is $\O_K$. We have the `obvious guess' of the order $\Z[\zeta]$. We need to check this guess. We know $\disc(\Z[\zeta])= (-1)^{\frac{(p-1)(p-2)}{2}} \Nm{K/\Q}(f'(\zeta))$. But what is $f'(\zeta)$? We first calculate $f'(x)$, then evaluate at $\zeta$:
	\[
	\begin{split}
	(x-1)f(x)&= x^p - 1 \\
	f(x) + (x-1) f'(x)&= p x^{p-1} \\
	f'(\zeta)&= \dfrac{p \,\zeta^{p-1}}{\zeta-1}
	\end{split}
	\]
Note that $[K \colon \Q]= \phi(p)=p-1$ and $\zeta \in \O_K^\times$. Let $\sigma_1,\ldots,\sigma_n$ be the complex embeddings of $K$ into $\C$. Then noting that $p-1$ is even and $f(x)=\prod_i (x-\sigma_i(\zeta))$, we have
	\[
	\begin{split}
	\Nm{K/\Q}(\zeta-1)&= (-1)^{p-1} \Nm{K/\Q}(1-\zeta) \\
	&=\Nm{K/\Q}(1-\zeta) \\
	&=\prod_{\sigma: K \hra \C} \sigma(1-\zeta) \\
	&= \prod_{\sigma: K \hra \C} (1-\sigma(\zeta)) \\
	&= f(1) \\
	&= 1 + 1 + \cdots + 1 \\
	&= p
	\end{split}
	\]
We know $(\zeta -1) f'(\zeta) = p \zeta^{p-1}$. Taking norms gives
	\[
	\begin{split}
	\Nm{K/\Q}(\zeta-1) \Nm{K/\Q}(f'(\zeta))&= \Nm{K/\Q}(p) \Nm{K/\Q}(\zeta)^{p-1} \\
	\Nm{K/\Q}(\zeta-1) \Nm{K/\Q}(f'(\zeta))&= p^{p-1} (\pm 1)^{p-1} \\
	p \Nm{K/\Q}(f'(\zeta))&= p^{p-1}
	\end{split}
	\]
Therefore, $\disc(\Z[\zeta])= (-1)^{\frac{(p-1)(p-2)}{2}} p^{p-2}=(-1)^{\frac{p-1}{2}} p^{p-2}$, since $p$ is odd. Note that we lose nothing by assuming that $p$ is odd: if $p=2$, then $\Q(\zeta)=\Q$. Then if $\Z[\zeta] \subsetneq \O_K$, it must be that $[\O_K \colon \Z[\zeta]]=p^e$ for some $e \geq 1$. After possibly multiplying by a power of $p$, we know there is a $\alpha \in \O_K$ such that $\alpha \notin \Z[\zeta]$ and $p\alpha \in \Z[\zeta]$. We know $\Z[\zeta]= \Z[\zeta-1]$. Then $\alpha$ can be written
	\[
	\alpha = \dfrac{a_0}{p} + \dfrac{a_1}{p} (\zeta-1) + \cdots + \dfrac{a_{p-2}}{p} (\zeta-1)^{p-2},
	\]
where $a_i \in \Z$ with not all $a_i$ divisible by $p$. Subtracting multiples of $p$ from those terms divisible by $p$, we may assume $a_i$ is not divisible by $p$ for $0 \leq i \leq p-2$. Hence, we can write
	\[
	\alpha = \dfrac{a_i}{p} (\zeta-1)^i + \cdots + \dfrac{a_{p-2}}{p} (\zeta-1)^{p-2},
	\]
where $p \nmid a_i$. Multiplying by $\frac{p}{(\zeta-1)^{i+1}}$, we have
	\[
	\dfrac{p\alpha}{(\zeta-1)^{i+1}} = \dfrac{a_1}{\zeta-1} + \underbrace{a_{i+1} + a_{i+2}(\zeta-1) + \cdots + a_{p-2}(\zeta-1)^{p-2-(i+1)}}_{\in\, \Z[\zeta]}
	\]
Now $\Nm{K/\Q}(\frac{a_i}{\zeta-1})= \frac{a_i^{p-1}}{p} \notin \Z$ as $p \nmid a_i$. But then $\frac{a_i}{\zeta-1} \notin \O_K$. But then using the above equality, this shows $\frac{p}{(\zeta-1)^{i+1}} \alpha \notin \O_K$. But as $\alpha \in \O_K$, this shows $\frac{p}{(\zeta-1)^{i+1}} \notin \O_K$. Therefore, $\frac{p}{(\zeta-1)^{i+1}} \notin \Z[\zeta]$. On the other hand,
	\[
	\begin{split}
	p&= \Nm{K/\Q}(\zeta-1) \\
	&= \prod_{\sigma} (\sigma(\zeta)-1) \\
	&=\prod_{i=1}^{p-1} (\zeta^i -1)
	\end{split}
	\]
Using the fact that $\zeta^i-1=(\zeta-1)(1+\zeta+\cdots+\zeta^{i-1})$, this shows that $(\zeta-1)^{p-1} \mid p$ in $\Z[\zeta]$. But as $i+1 \leq p-1$, this contradicts the fact that $\frac{p}{(\zeta-1)^{i+1}} \notin \Z[\zeta]$. Therefore, we know 
	\[
	\begin{split}
	\O_K&=\Z[\zeta] \\
	\disc \O_K&= (-1)^{\frac{p-1}{2}} p^{p-2}
	\end{split}
	\]
Note that $\frac{p}{(\zeta-1)^{p-1}} \in \Z[\zeta]$ and $\Nm{K/\Q}(\frac{p}{(\zeta-1)^{p-1}})= \frac{p^{p-1}}{p^{p-1}}=1$ so that $p=u(\zeta-1)^{p-1}$ for some $u \in \Z[\zeta]^\times$. \xqed
\end{ex}

\begin{rem}
Choose $m \geq 1$ and let $\zeta_m$ be the primitive $m$th root of unity. Then $\O_K=\Z[\zeta_m]$ and $\disc \O_K \mid m^{\phi(m)}$. 
\end{rem}



