% !TEX root = ../../math6370.tex
\section{Introduction}

\noindent\textbf{MATH 6370:} An introduction to number theory focusing on the algebraic theory. Topics include, but are not limited to: number fields, Dedekind domains, class groups, Dirichlet's unit theorem, local fields, ramification, decomposition and inertia groups, and the distribution of primes. \\

\noindent \textbf{Recommended Text References}

\begin{itemize}
\item Marcus, \emph{Number Fields}. Apparently good for beginners and has lots of exercises. Chapter 1 is motivational and can be skipped if desired. It avoids local fields and Dedekind domains.

\item Neukirch, \emph{Algebraic Number Theory}. This text is more advanced and treats the subject from the general point of view of arithmetic geometry (which may seem strange to those without the geometric background).

\item Milne, \emph{Algebraic Number Theory}. Milne's course notes (in several subjects) are always good and freely available.

\item Lang, \emph{Algebraic Number Theory}

\item Murty, Esmonda, \emph{Problems in Algebraic Number Theory}. This book was designed for self study. Lots of exercises with full solutions.

\item Janusz, \emph{Algebraic Number Fields}
\end{itemize}

These notes were taken in Spring 2018 in a course taught by Professor David Zywina at Cornell University. In some places, notation/material has been changed or added. Any errors in this text should be attributed to the typist --- Caleb McWhorter --- and not the instructor or any referenced text. 